---
title: <center>Gene and Species Trees</center>
author: <center></center>
date: <center>`r format(Sys.Date(),format="%B %d %Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---

Trying to write some basic code to do gene and species trees.  if it works, we'll try to write it up.

#### Basic ms call

Situation we have is two populations, with overall &theta; = 11, which have been separate for 10,000 generations.  First, we'll just look at a few trees:

```{r}
dat.1 <-ms(nsam=10,nreps=1,opts=c("-t 10 -I 2 5 5  -ej .375 2 1 -T"))

tre.1 <-read.tree(text=dat.1)

nms <-tre.1$tip.label
nm2 <-as.numeric(sapply (nms, function (x) substr(x,2,nchar(x))))
nmcol=nm2
nmcol[which(nm2<=5)]="red"
nmcol[which(nm2>5&nm2<=20)]="darkgreen"
#nmcol[which(nm2>20]="blue" 
tr <-plot(tre.1,tip.color=nmcol)
# abline(v=tr$x.lim[2]-.375,col="red")
axisPhylo()
#add.scale.bar(cex = 0.7, font = 2, col = "red")
tr$x.lim
```
Need to work on this.  Maybe not worth it.


Interesting - so with the above, in the absence of migration, lineage sorting always occurs.  Actually, that sort of makes sense.  But enough for now.


#### 1.  How about simple gene geneaology

The following has no mutation
```{r}
ms1 <-ms(nsam=10,nreps=1, opts=c("-T"))
tree1 <-read.tree(text=ms1)
plot(tree1,edge.width=2)
axisPhylo()
```

Now add some random mutations (by adding theta)

```{r}
ms2 <-ms(nsam=10,nreps=1, opts= c("-t 30 -T"))
tree2 <-read.tree(text=ms2)
plot(tree2,edge.width=2)
axisPhylo()
```
Not too much different, right?

#### Throw in some expansion.

Basic code taken from SkylinePlots.Rmd

```{r}
N <-100000
r <-.02 #growth rate per year
gen <-5 #generations per year
nyears <-(4*N)/5 # 1 coalescent unit in years
g <-r*nyears # growth rate in coalescence units
m.out <-ms(nsam=20, nreps=1, opts=(paste("-t 12 -T -r 0 702 -G",g,sep=" ")))
tree.gr <-read.tree(text=m.out)
plot(tree.gr,edge.width=2,cex=.8)
axisPhylo()
```
```{r}
plot(skyline(tree.gr,epsilon=.0004),lwd=2)
```
So we see that it leads to longer branches - need to explain.

#### Bottleneck

Now use modification of example from ms documentation

```{r}
ms3 <-ms(nsam=30,nreps=1,opts=c("-t 12 -G 6.93 -eG 0.2 0.0 -eN 0.3 0.5 -T"))
tree3 <-read.tree(text=ms3)
plot(tree3)
axisPhylo()
```
And do a skyline
```{r}
plot(skyline(tree3,epsilon=.0015))
```

So we'll think about all of this.  