---
title: <center>Some Tree Building</center>
author: <center></center>
date: <center>`r format(Sys.Date(),format="%%B %%d %%Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---

### Background

For the moment, this is just a scratch file, but we're going to see if we can use stuff from summer to make trees complete with the Geo time scale.

```{r}
library (strap)
```

Basic code is something like

geoscalePhylo(tree=ladderize(file.tree,right=FALSE),label.offset=0,cex.age=.6,cex.ts=.6,cex.tip=.6)

Where file.tree is a phylo object.  Note also that tree.file$root.time  may also have to be set with max(dist.nodes(tree.file))

So where to get the trees?  Probably start with treebase

```{r}
library(treebase)
```
Let's try a simple query
```{r}
humans <-search_treebase("Homo",by="taxon")
```
In actuality, it might make sense to do the search on the web and then select specific one to download.  Ok< I have one for Carnivores (StudyID =Tr3367)
```{r}
car.tree <-read.nexus("./Data/Carnivora.nex")
```
```{r}
plot(car.tree)
```
OK.  It works out pretty nicely, but it doesn't look to have branch lengths.  Let's explore
```{r}
str(car.tree)
```
Nope.  No dates.  But we can explore one thing.  So the taxa are
```{r}
car.tree$tip.label
```
So I wonder if we can get a list of common names from that
```{r}
library(taxize)
```
```{r}
car.nm <-sci2comm(car.tree$tip.label,simplify = TRUE)
```
Returns a list, with lots of synonyms for each one.  Adding simplify=FALSE gives more information, but there's no real way to simplify things.

Back to taxize.  There's a study, S1663, of placental mammals
```{r}
mammals <-search_treebase("S1663",by="id.study")
```
```{r}
str(mammals)
```
So that returned three trees, all way too big for practical use.  but we could walk through the code anyway, just to see if calibration works.  Pick the first one

```{r}
mammal1 <-mammals[[1]]
```
```{r}
mammal1$root.time <-max(dist.nodes(mammal1))
mammal1$root.time
```
Hmm.

Go back to downloaded file
```{r}
cats <-read.nexus("./Data/Felid.nex")
plot(cats)
```
Lotsof multichotomies.  Not good.

Found another source - lots and lots of dated trees - reference to follow.  From Europe.
```{r}
emam <- read.tree("./Data/EMammals.txt")
```
Ok.  it's way to large.  However, We should be able to do the manipulations:
```{r}
emam$root.time <-max(dist.nodes(emam))
emam$root.time
```
Looks quite good.  Now to do the stuff
```{r}
geoscalePhylo(tree=ladderize(emam,right=FALSE),label.offset=0,cex.age=.6,cex.ts=.6,cex.tip=.2,units="Period")
```
OK.  So now we get to the pruning piece.  First, list the taxa
```{r}
emam$tip.label
```
```{r}
test <-read.tree("./Data/newick.txt")
plot(test)
```
OK.  So the trick is to get the topology correct in the Newick file and then add the numbers.

```{r}
test$root.time <-max(dist.nodes(test))
geoscalePhylo(tree=ladderize(test,right=FALSE),label.offset=0,cex.age=.6,cex.ts=.6,cex.tip=.8,units="Period")
```

